---
title: "Do penguins with longer flippers weigh more?"
author: "Alice"
date: "2026-09-21"
engine: knitr
---

## Question

How does flipper length relate to body mass, and does the relationship look similar across penguin species?

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/), provided through the R `palmerpenguins` package.

## Load and inspect the data

```{r}
#| label: load-r-data
#| message: false
#| warning: false

library(palmerpenguins)
library(ggplot2)

head(penguins)
summary(penguins[c("flipper_length_mm", "body_mass_g")])
```

## Prepare the measurements

I remove rows missing either measurement so that each plotted penguin has both a flipper length and a body mass.

```{r}
#| label: prepare-r-data

keep <- complete.cases(penguins[c("flipper_length_mm", "body_mass_g")])
penguins_clean <- penguins[keep, ]

data.frame(
    original_rows = nrow(penguins),
    retained_rows = nrow(penguins_clean),
    removed_rows = sum(!keep)
)
```

## Plot the relationship

```{r}
#| label: fig-flipper-mass
#| fig-cap: "Penguin body mass versus flipper length, coloured by species."
#| echo: true

ggplot(
    penguins_clean,
    aes(x = flipper_length_mm, y = body_mass_g, colour = species)
) +
  geom_point(alpha = 0.7) +
  labs(
    x = "Flipper length (mm)",
    y = "Body mass (g)",
    colour = "Species"
  ) +
  theme_minimal()
```

## Findings

Penguins with longer flippers generally have greater body mass in this sample. Gentoo penguins tend to have longer flippers and greater body mass than Adelie and Chinstrap penguins.

The overall pattern partly reflects differences between species.
This plot shows an association; it does not establish that longer flippers cause greater body mass.