---
title: "Penguin body mass with Python and R"
author: "Alice"
date: "2026-09-24"
engine: knitr
---
```{r}
#| label: setup
#| message: false
#| warning: false

library(reticulate)
```

```{python}
#| label: check-python

import os
import sys
from pathlib import Path

project_venv = Path(os.environ["QUARTO_PROJECT_DIR"]) / ".venv"

print("Python executable:", sys.executable)
print("Using project .venv:", os.path.samefile(sys.prefix, project_venv))
```

 ## Comparing penguin body mass

I use Python to calculate average body mass for each penguin species,
then pass the summary to R to visualize the comparison.

Data: [Palmer Penguins](https://allisonhorst.github.io/palmerpenguins/),
provided through the Python `palmerpenguins` package.

Then add a Python chunk to load the data:

```{python}
#| label: load-penguins

from palmerpenguins import load_penguins

penguins = load_penguins()
penguins.head()
```

penguins stores the dataset. .head() displays the first five rows so you can inspect its columns.

Next, add a second Python chunk:

```{python}
#| label: summarize-mass

penguins_clean = penguins.dropna(subset=["species", "body_mass_g"])

mass_summary = (
    penguins_clean.groupby("species", as_index=False)
    .agg(mean_mass_g=("body_mass_g", "mean"))
)

mass_summary
```

Here’s what it does:

- dropna(...) excludes rows missing the values needed for this comparison.
- groupby(...) groups penguins by species. as_index=False keeps species as a regular column, convenient for plotting
    in R.

- .agg(...) calculates mean body mass and names the result column mean_mass_g.

## Visualizing the summary in R

I pass the summary calculated in Python to R using reticulate.

```{r}
#| label: transfer-summary

mass_summary_r <- py$mass_summary
mass_summary_r
```

This is the key bonus step:

- py$mass_summary accesses the object you created in Python.
- Reticulate converts the Python pandas DataFrame into an R data frame.
- <- assigns it to the R object mass_summary_r.

Now add a plot using that transferred data:

```{r}
#| label: fig-mass-python-r
#| fig-cap: "Mean penguin body mass by species, calculated in Python and plotted in R."

library(ggplot2)

ggplot(mass_summary_r, aes(x = species, y = mean_mass_g)) +
geom_col() +
labs(
    x = "Species",
    y = "Mean body mass (g)"
) +
theme_minimal()
```

geom_col() draws bars using your already-calculated means. R is using Python’s result here, which demonstrates the
required object transfer.


Gentoo penguins had the highest average body mass in this dataset, at approximately 5,076 g. Adelie and Chinstrap
penguins had similar averages, at about 3,701 g and 3,733 g, respectively. These means summarize differences between
species but do not show variation among individual penguins or how much their body masses overlap.